### 2021-02-06 manuscript/manuscript.Rmd
Use original data instead of cleaned

---
title: "Results MSRC latent class analysis"
output: github_document
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: references.bib
knit: worcs::cite_all
---

```{r setup, include=FALSE}
library("worcs")
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
load_data()
knitr::opts_chunk$set(echo = TRUE)
```

This manuscript uses the Workflow for Open Reproducible Code in Science [@vanlissaWORCSWorkflowOpen2020] to ensure reproducibility and transparency. All code <!--and data--> are available at <https://github.com/cjvanlissa/MSRC_LPA>.

<!--The function below inserts a notification if the manuscript is knit using synthetic data. Make sure to insert it after load_data().-->
`r notify_synthetic()`

## Descriptive statistics

```{r}
library(ggplot2)
library(tidySEM)
library(tidyLPA)
desc <- descriptives(df)

use_vars <- c("CDE7_NEW", "CDE8", "CDE9", "CDE10", "CDE11", "CDE12", "SA_Freq_Rank", "NSSI_Freq_Rank")

df_plot <- do.call(rbind, lapply(grep("^CD", use_vars, value = TRUE), function(x){data.frame(Variable = x, value = df[[x]])}))

p_bar <- ggplot(df_plot, aes(x = value)) + 
  geom_bar() +
  facet_wrap(~Variable, scales = "free") +
  theme_bw()

ggsave("bar.png", device = "png", width = 200, height = 120, units = "mm", dpi = 300)

df_plot <- do.call(rbind, lapply(grep("Freq", use_vars, value = TRUE), function(x){data.frame(Variable = x, value = df[[x]])}))

df$NSSI_log <- log(df$NSSI_Freq_Rank+.0000001)
df$SA_log <- log(df$SA_Freq_Rank+.0000001)
df$NSSI_inv <- 1/(df$NSSI_Freq_Rank+.0000001)
df$SA_inv <- 1/(df$SA_Freq_Rank+.0000001)
df$NSSI_sqrt <- sqrt(df$NSSI_Freq_Rank)
df$SA_sqrt <- sqrt(df$SA_Freq_Rank)
df$NSSI_dummy <- as.numeric(df$NSSI_Freq_Rank == 0)
df$SA_dummy <- as.numeric(df$SA_Freq_Rank == 0)
df_plot <- do.call(rbind, lapply(grep("(NSSI|SA)", names(df), value = TRUE), function(x){
  data.frame(Variable = paste0("Variable: ", gsub("^(.+?)_.*$", "\\1", x), " (", gsub("_", " ", gsub("^.+?_(.*)$", "\\1", x)), ")"), 
             value = df[[x]])
}))
df_plot$Variable <- ordered(df_plot$Variable, levels = unique(df_plot$Variable))
p_dens <- ggplot(df_plot, aes(x = value)) + 
  geom_density() +
  facet_wrap(~Variable, scales = "free", ncol = 2) +
  theme_bw()

ggsave("density.png", device = "png", width = 200, height = 350, units = "mm", dpi = 300)
```

## Analysis

```{r model}

# MSRC <- data[1:1905,] %>%
#   select (CDE7_NEW, CDE8, CDE9, CDE10, CDE11, CDE12, SA_Freq_Rank, NSSI_Freq_Rank) %>%
#   single_imputation() %>%
#   estimate_profiles(1:10, variances="equal", covariances="zero")
# 
# MSRC
# 
# LPABest <- data[1:1905,] %>%
#   select(CDE7_NEW, CDE8, CDE9, CDE10, CDE11, CDE12, SA_Freq_Rank, NSSI_Freq_Rank) %>%
#   single_imputation() %>%
#   estimate_profiles(9, variances="equal", covariances="zero")
# get_fit(LPABest)
# 
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.